Statement of Intent

### 2010-11-17 applications/nsf_fellowship/personal_statement.tex
I'm pretty happy with my personal statement right now, which surprises me.  I didn't think it was that good at first.

I came back from my experience in Alabama with a new-found desire to study
computer science.  I changed my major to computer science, with minors in
physics and mathematics, as those subjects were still interesting to me and
they are relevant background for a computer scientist.  I found a computer
science professor who was doing research that I found intriguing; I wanted to
learn about machine translation, which I knew was part of machine learning, so
I started working in the Applied Machine Learning Lab with Dr. Kevin Seppi.

My time working in the Applied Machine Learning lab has been the defining
experience of my undergraduate education.  I had very close interaction with
several professors and many graduate students who showed me how interesting
research can be.  From those that I had association with, I gained a love of
learning new things and exploring new ideas.  I also learned a great deal more
than was taught in my classes in order to understand the research that was
going on in the lab.  Several times in my classes the professor teaching the
class had me give part of a lecture that dealt with things I was actively
researching.

@@

I also had a great internship experience as an undergraduate that helped to
shape my vision of what I wanted from my career.  I worked as an intern on the
TV Ads team at Google during the summer of 2009.  My intern host didn't have a
clear vision of the project she wanted me to do, so I proposed a meaningful,
difficult project that, if it worked, would have a large impact on the
company's business.  My host supported me and gave me the help that I needed,
and the project turned out to be very successful.  One of the main things I
learned from the experience was that while I enjoy writing software, I find
much more satisfication in solving interesting problems that actually improve
people's lives.

As I was nearing the end of my undergraduate career, I found that I had several
research projects that needed just a little more time to be finished.  I wanted
to get into one of the nation's top graduate schools for a PhD, and I thought
that having several publications, instead of hopes of publishing, would give me
a much stronger application.  I also had taken several graduate-level classes
as an undergraduate that wouldn't transfer well to another university, and I
was really enjoying working with Dr. Seppi.  All of these pointed me towards a
masters degree at BYU, so I decided to stay, hoping to finish within a year.

My first semester as a masters student during the summer of 2010 gave me my
first real teaching opportunity.  While I have tutored other students, mostly
informally, throughout all of my college career, I had never done so in a
formal setting.  That semester, however, I was a teaching assistant for two
classes that Dr. Seppi taught: Introduction to Articial Intelligence, and
Bayesian Inference.  I had always enjoyed helping people understand concepts,
but I didn't realize how much I would like the teaching aspect of being a
professor until that experience.  I found that figuring out how to explain
complicated concepts simply and seeing the light of understanding in the
students' faces gave me great satisfaction.

>>>>

I came back from my experience in Alabama with a new-found desire to study
computer science.  I changed my major to computer science, with minors in
physics and mathematics, as those subjects were still interesting to me and
they are relevant background for a computer scientist.  I found a computer
science professor who was doing research that I found intriguing; I wanted to
learn about machine translation, which I knew was part of machine learning, so
I started working in the Applied Machine Learning lab with Dr. Kevin Seppi.

My time working in the Applied Machine Learning lab has been the defining
experience of my undergraduate education.  I had very close interaction with
several professors and many graduate students who showed me how interesting
research can be.  From those that I had association with I gained a love of
learning new things and exploring new ideas.  I also learned a great deal more
than was taught in my classes in order to understand the research that was
going on in the lab.  Several times in my classes the professor teaching the
class had me give part of a lecture that dealt with things I was actively
researching.

@@

I also had a great internship experience as an undergraduate that helped to
shape my vision of what I want from my career.  I worked as an intern on the TV
Ads team at Google during the summer of 2009.  My intern host didn't have a
clear idea for the project she wanted me to do, so I proposed a meaningful,
difficult project that, if it worked, would have a large impact on the
company's business.  My host supported me and gave me the help that I needed,
and the project turned out to be very successful.  The project was so
successful that the director of all of TV Ads at Google said that my project
was the most important work going on in the whole TV Ads team, and several
times he commended me for my work.  One of the main things I learned from the
experience was that while I enjoy writing software, I find much more
satisfication in solving interesting problems that actually improve people's
lives.  I was able to help the TV Ads team, which was struggling to make money,
greatly improve their business.

I returned from my internship with one year left as an undergraduate.  As that
year neared its end and the time came to apply to graduate schools, I found
that I had several research projects that needed just a little more time to be
finished.  I wanted to finish what I had started here at BYU.  I also wanted to
get into one of the nation's top universities for a PhD, and I thought that
having several publications, instead of just hopes of publishing, would give me
a much stronger application.  Further, I had taken several graduate-level
classes as an undergraduate that wouldn't transfer well to another university,
and I was really enjoying working with Dr. Seppi.  All of these pointed me
towards a masters degree at BYU, so I decided to stay, hoping to finish within
a year.

My first semester as a masters student during the summer of 2010 gave me my
first real teaching opportunity.  While I have tutored other students
throughout all of my college career, I had never done so in a formal setting.
That semester, however, I was a teaching assistant for two classes that Dr.
Seppi taught: Introduction to Articial Intelligence, and Bayesian Inference.  I
had always enjoyed helping people understand concepts, but I didn't realize how
much I would like the teaching aspect of being a professor until that
experience.  I found that figuring out how to explain complicated concepts
simply and seeing the light of understanding in the students' faces gave me
great satisfaction.
